IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2020-10-03 main.tex
:ambulance: Fix LoS and NLoS tap gains
:pencil: Add plots

\begin{document}
	\title{IRS-aided SWIPT: Joint waveform, active and passive beamforming design}
	\author{
		\IEEEauthorblockN{
			Yang~Zhao,~\IEEEmembership{Student~Member,~IEEE,}
			Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE,}
			and~Zhenyuan~Feng,~\IEEEmembership{Student~Member,~IEEE}
		}
		\thanks{
			The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: yang.zhao18@imperial.ac.uk; b.clerckx@imperial.ac.uk).
		}
	}
	\maketitle

@@

			By discarding the non-DC components, taking expectation over symbol distribution, and truncating \ref{eq:i_d} to the $n_0$-th order, we approximate the average output DC current for a given channel as
			\begin{equation}\label{eq:i_out}
				i_{\text{out}}(t)=\mathcal{A}\{i_d(t)\}\approx\sum_{i=0}^{\infty}{k_i'}{\rho^{i/2}}{R_{\text{ant}}^{i/2}}\mathcal{E}\left\{{\mathcal{A}\left\{y(t)^i\right\}}\right\}
			\end{equation}

@@

			When $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, the optimal phase shift vector $\bar{\boldsymbol{\phi}}^\star$ can be obtained by Eigenvalue Decomposition (EVD). Otherwise, a suboptimal solution can be extracted via Gaussian randomization method \cite{Huang2010}. Specifically, we perform EVD $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$, generate $Q$ CSCG random vectors $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1}),\ q=1,\dots,Q$, construct the corresponding candidates $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)}$, and choose the one that maximizes the objective function \ref{ob:irs}. Finally, the phase shift is retrieved by $\theta_l=\arg(\phi_l^\star/\phi_{L+1}^\star), \ l=1,\dots,L$. The algorithm for phase shift optimization is summarized in Algorithm~\ref{al:irs}.
			\begin{algorithm}[!t]
				\caption{SCA: IRS Phase Shift}
				\label{al:irs}
				\begin{algorithmic}[1]
					\State \textbf{input} $\beta_2,\beta_4,\boldsymbol{h}_{D,n},\boldsymbol{H}_{I,n},\boldsymbol{h}_{R,n},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho,\sigma_n,\bar{R},Q,\epsilon$
					\State Construct $\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n}$ for $n=1,\dots,N$, $\boldsymbol{C}_{I/P,n}$ for $n=-N+1,\dots,N-1$
					\State \textbf{initialize} $i \gets 0,\boldsymbol{\Phi}^{(0)},t_{I/P,n}^{(0)}$ for $n=-N+1,\dots,N-1$
					\Repeat
						\State $i \gets i + 1$
						\State Obtain $\boldsymbol{\Phi}^{(i)}, t_{I/P,n}^{(i)}$ by solving problem~\ref{op:irs}
						\State Compute $z^{(i)}$ by \ref{eq:z_irs}
					\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
					\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Obtain $\bar{\boldsymbol{\phi}}^\star$ by EVD, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Obtain $\boldsymbol{U},\boldsymbol{\Sigma}$ by EVD, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$

						\State Generate $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $q=1,\dots,Q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)}$, $\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$
						\State Set $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)}$, $\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\phi_l^\star/\phi_{L+1}^\star), \ l=1,\dots,L$, construct $\boldsymbol{\phi}^{\star}$
					\State \textbf{output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}
		\end{subsection}

@@

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs.tex}
			\end{adjustbox}
			\caption{Average R-E region for adaptive, ideal, fixed and no IRS over $B=10\,\si{\MHz}$}
			\label{fi:re_irs}
		\end{figure}

		Fig.~\ref{fi:re_irs} and [TODO] explore the average R-E region under different IRS configuration for narrowband transmission ($B=1\,\si{\MHz}$) and broadband transmission ($B=10\,\si{\MHz}$). The adaptive scheme optimizes the IRS and waveform alternatively for each points in the R-E boundary. In comparison, the WIT/WPT-based schemes only perform alternating optimization for the right-most/left-most points, then fix the IRS and update the waveform to obtain the R-E curve. To gain some insight into the IRS behavior, we compare the results above to that of no IRS and the ideal FS IRS, where we assume each element has independent and adjustable reflection coefficients over all subbands such that the ideal IRS has a total DoF of $NL$. Since the IRS only adapts the phase of the extra channel, the optimal strategy for each FS reflector in the SISO case would be aligning the AP-IRS-user and AP-user channel over all subbands, namely
		\begin{equation}
			\theta_{l,n}^{\star}=e^{j\arg{\left(h_{D,n}/(h_{I,n,l}h_{R,n,l})\right)}}
		\end{equation}
		\textit{First}, it is observed that the presence of IRS effectively enlarges the achievable R-E region in both cases. This is because the IRS tweaks the weak extra channels such that they add constructively to enhance the composite channel. \textit{Second}, the performance gaps of the adaptive, ideal and fixed IRS are negligible for narrowband transmission but noticeable for broadband transmission. The reason is that when $B=1\,\si{\MHz}$, all channels are approximately flat and the response of all subbands are roughly the same. In such cases, the additional DoF provided by the frequency selectivity of the ideal IRS would be unnecessary as the the optimal reflection coefficients of each element would almost align at all subbands. Similarly, WIT-optimized and WPT-optimized IRS boil down to optimizing a single term representing the composite channel response at all subbands. Therefore, all IRS strategies coincide with each other, and the optimal IRS for narrowband SISO SWIPT can be approximated by any candidate that roughly align the AP-IRS-user channel with the AP-user channel simultaneously over all subbands. On the other hand, the channel frequency selectivity becomes significant for $B=10\,\si{\MHz}$, and the ideal FS IRS outperforms the others as it requires no tradeoff among subchannels. Moreover, the WIT-optimized IRS tends to equalize the channel strength for all subbands as possible, due to the preference of WF strategy at high SNR. In comparison, the WPT-optimized IRS beams less towards few weakest subchannels, due to the inefficiency of small-amplitude tones in energy harvesting. Therefore, the adaptive IRS design is more suitable for broadband SWIPT. Note that the tradeoff for practical IRS indeed origins from frequency selectivity rather than number of subbands.
	\end{section}

>>>>

\begin{document}
	\title{IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design}
	\author{
		\IEEEauthorblockN{
			Yang~Zhao,~\IEEEmembership{Student~Member,~IEEE,}
			Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE,}
			and~Zhenyuan~Feng,~\IEEEmembership{Student~Member,~IEEE}
		}
		\thanks{
			The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: yang.zhao18@imperial.ac.uk; b.clerckx@imperial.ac.uk).
		}
	}
	\maketitle

@@

			By discarding the non-DC components, taking expectation over symbol distribution, and truncating \ref{eq:i_d} to the $n_0$-th order, we approximate the average output DC current for a given channel as
			\begin{equation}\label{eq:i_out}
				i_{\text{out}}(t)=\mathcal{A}\{i_d(t)\}\approx\sum_{i=0}^{n_0}{k_i'}{\rho^{i/2}}{R_{\text{ant}}^{i/2}}\mathcal{E}\left\{{\mathcal{A}\left\{y(t)^i\right\}}\right\}
			\end{equation}

@@

			When $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, the optimal phase shift vector $\bar{\boldsymbol{\phi}}^\star$ can be obtained by Eigenvalue Decomposition (EVD). Otherwise, a suboptimal solution can be extracted via Gaussian randomization method \cite{Huang2010}. Specifically, we perform EVD $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$, generate $Q$ CSCG random vectors $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1}),\ q=1,\dots,Q$, construct the corresponding candidates $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)}$, and choose the one that maximizes the objective function \ref{ob:irs}. Finally, the phase shift is retrieved by $\theta_l=\arg(\phi_l^\star/\phi_{L+1}^\star), \ l=1,\dots,L$. The algorithm for phase shift optimization is summarized in Algorithm~\ref{al:irs}.
			\begin{algorithm}[!t]
				\caption{SCA: IRS Phase Shift}
				\label{al:irs}
				\begin{algorithmic}[1]
					\State \textbf{input} $\beta_2,\beta_4,\boldsymbol{h}_{D,n},\boldsymbol{H}_{I,n},\boldsymbol{h}_{R,n},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho,\sigma_n,\bar{R},Q,\epsilon$
					\State Construct $\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n}$ for $n=1,\dots,N$, $\boldsymbol{C}_{I/P,n}$ for $n=-N+1,\dots,N-1$
					\State \textbf{initialize} $i \gets 0,\boldsymbol{\Phi}^{(0)},t_{I/P,n}^{(0)}$ for $n=-N+1,\dots,N-1$
					\Repeat
						\State $i \gets i + 1$
						\State Obtain $\boldsymbol{\Phi}^{(i)}, t_{I/P,n}^{(i)}$ by solving problem~\ref{op:irs}
						\State Compute $z^{(i)}$ by \ref{eq:z_irs}
					\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
					\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Obtain $\bar{\boldsymbol{\phi}}^\star$ by EVD, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Obtain $\boldsymbol{U},\boldsymbol{\Sigma}$ by EVD, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$
						\State Generate $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $q=1,\dots,Q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)}$, $\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$
						\State Set $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)}$, $\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\phi_l^\star/\phi_{L+1}^\star), \ l=1,\dots,L$, construct $\boldsymbol{\phi}^{\star}$
					\State \textbf{output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}
		\end{subsection}

@@

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs_1mhz.tex}
			\end{adjustbox}
			\caption{Average R-E region for adaptive, ideal, fixed and no IRS over $B=1\,\si{\MHz}$}
			\label{fi:re_irs_1mhz}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs_10mhz.tex}
			\end{adjustbox}
			\caption{Average R-E region for adaptive, ideal, fixed and no IRS over $B=10\,\si{\MHz}$}
			\label{fi:re_irs_10mhz}
		\end{figure}

		Fig.~\ref{fi:re_irs_1mhz} and \ref{fi:re_irs_10mhz} explore the average R-E region under different IRS configuration for narrowband transmission ($B=1\,\si{\MHz}$) and broadband transmission ($B=10\,\si{\MHz}$). The adaptive scheme optimizes the IRS and waveform alternatively for each points in the R-E boundary. In comparison, the WIT/WPT-based schemes only perform alternating optimization for the right-most/left-most points, then fix the IRS and update the waveform to obtain the R-E curve. To gain some insight into the IRS behavior, we compare the results above to that of no IRS and the ideal FS IRS, where we assume each element has independent and adjustable reflection coefficients over all subbands such that the ideal IRS has a total DoF of $NL$. Since the IRS only adapts the phase of the extra channel, the optimal strategy for each FS reflector in the SISO case would be aligning the AP-IRS-user and AP-user channel over all subbands, namely
		\begin{equation}
			\theta_{l,n}^{\star}=e^{j\arg{\left(h_{D,n}/(h_{I,n,l}h_{R,n,l})\right)}}
		\end{equation}
		\textit{First}, it is observed that the presence of IRS effectively enlarges the achievable R-E region in both cases. This is because the IRS tweaks the weak extra channels such that they add constructively to enhance the composite channel. \textit{Second}, the performance gaps of the adaptive, ideal and fixed IRS are negligible for narrowband transmission but noticeable for broadband transmission. The reason is that when $B=1\,\si{\MHz}$, all channels are approximately flat and the response of all subbands are roughly the same. In such cases, the additional DoF provided by the frequency selectivity of the ideal IRS would be unnecessary as the the optimal reflection coefficients of each element would almost align at all subbands. Similarly, WIT-optimized and WPT-optimized IRS boil down to optimizing a single term representing the composite channel response at all subbands. Therefore, all IRS strategies coincide with each other, and the optimal IRS for narrowband SISO SWIPT can be approximated by any candidate that roughly align the AP-IRS-user channel with the AP-user channel simultaneously over all subbands. On the other hand, the channel frequency selectivity becomes significant for $B=10\,\si{\MHz}$, and the ideal FS IRS outperforms the others as it requires no tradeoff among subchannels. Moreover, the WIT-optimized IRS tends to equalize the channel strength for all subbands as possible, due to the preference of WF strategy at high SNR. In comparison, the WPT-optimized IRS beams less towards few weakest subchannels, due to the inefficiency of small-amplitude tones in energy harvesting. Therefore, the adaptive IRS design is more suitable for broadband SWIPT. Note that the tradeoff for practical IRS indeed origins from frequency selectivity rather than number of subbands.
	\end{section}
